---
title: "Day XX Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(igraph)
library(ggraph)
library(readr)
library(collections)
options(scipen = 999)
```


## Part 1

```{r}
input <- read_lines("Day21Sample.txt")

```

Going to be lazy and just run it through until root has a value

```{r}
rootfinder<-function(numberyell=input){
  done<-FALSE
  ### set up a dictionary
  mnk<-dict()
  while(done==FALSE){
    ### run through all of the lines of input
    i<-1
    while(i<=length(numberyell)){
      ### pull apart the line into "words"
      yell<-unlist(str_split(str_replace_all(numberyell[i],":","")," "))
      ### if only two things, then
      if(length(yell)==2){
        ### add it to the dictionary
        mnk$set(key=yell[1],value=as.numeric(yell[2]))
        ###remove it from the list because there's no reason to re-run this.
        numberyell<-numberyell[-i]
        ## check if the monkey has enough information to do anything
      }else{if(mnk$has(yell[2])&&mnk$has(yell[4])){
        ### if so find its value and remove it from the list
        switch(yell[3],
               "+"=mnk$set(key=yell[1],value=mnk$get(yell[2])+mnk$get(yell[4])),
               "-"=mnk$set(key=yell[1],value=mnk$get(yell[2])-mnk$get(yell[4])),
               "*"=mnk$set(key=yell[1],value=mnk$get(yell[2])*mnk$get(yell[4])),
               "/"=mnk$set(key=yell[1],value=mnk$get(yell[2])/mnk$get(yell[4])),
               cat("no operator\n"))
        numberyell<-numberyell[-i]
        ### otherwise, increase i & do the next line
      }else{i<-i+1}}
      if(mnk$has("root")){done<-TRUE}}}
    mnk$get("root")}

```

```{r}
part1<-rootfinder()
part1
```
## Part 2
I'm going to try just searching for it

```{r}
### remove humn
nohuman<-input[-which(str_detect(input,"humn:"))]
### because I know root is a plus, change it to a minus
nohuman[which(str_detect(nohuman,"root:"))]<-str_replace(nohuman[which(str_detect(nohuman,"root:"))],"\\+","-")



humanfinder<-function(numberyell=input,human){
  done<-FALSE
  ### set up a dictionary
  mnk<-dict()
  mnk$set(key="humn",value=human)
  while(done==FALSE){
    ### run through all of the lines of input
    i<-1
    while(i<=length(numberyell)){
      ### pull apart the line into "words"
      yell<-unlist(str_split(str_replace_all(numberyell[i],":","")," "))
      ### if only two things, then
      if(length(yell)==2){
        ### add it to the dictionary
        mnk$set(key=yell[1],value=as.numeric(yell[2]))
        ###remove it from the list because there's no reason to re-run this.
        numberyell<-numberyell[-i]
        ## check if the monkey has enough information to do anything
      }else{if(mnk$has(yell[2])&&mnk$has(yell[4])){
        ### if so find its value and remove it from the list
        switch(yell[3],
               "+"=mnk$set(key=yell[1],value=mnk$get(yell[2])+mnk$get(yell[4])),
               "-"=mnk$set(key=yell[1],value=mnk$get(yell[2])-mnk$get(yell[4])),
               "*"=mnk$set(key=yell[1],value=mnk$get(yell[2])*mnk$get(yell[4])),
               "/"=mnk$set(key=yell[1],value=mnk$get(yell[2])/mnk$get(yell[4])),
               cat("no operator\n"))
        numberyell<-numberyell[-i]
        ### otherwise, increase i & do the next line
      }else{i<-i+1}}
      if(mnk$has("root")){done<-TRUE}}}
    mnk$get("root")}
```


try running this with powers of 10 to see if it approaches (or doesn't, 0)

```{r}
i<-0:15
i<-10^i

sapply(i,function(x){humanfinder(nohuman,x)})

```
based on this observation, I'm going to assume that if the number is too high, root is positive, if it is too low, root is negative. This is a huge assumption but if true, then it can be narrowed down to the right number.

Also, that the answer is less than 10^15

```{r}
human<-0
for(i in 15:0){
  try<-(0:9*10^i)+human
  cat(try,"\n")
  outs<-sapply(try,function(x){humanfinder(nohuman,x)})
  cat(outs,"\n")
  x<-max(which(outs<=0))
#  x<-max(which(outs>=0))  ### needed for real data.
  human<-try[x]
  cat("human = ",human,"\n")}

part2<-human
part2
```


