---
title: "Day XX Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(igraph)
library(ggplot2)
library(rgl)
library(ggraph)
library(readr)
library(ds4psy)
library(numbers)
options(scipen = 999)
```


```{r}
input<-read_lines("Day17Sample.txt")
jets<-unlist(str_split(input,""))
```

Set up the shapes

```{r}
shapes<-list(list(c(0,0),c(1,0),c(2,0),c(3,0)), # -
list(c(1,0),c(0,1),c(1,1),c(2,1),c(1,2)), # +
list(c(0,0),c(1,0),c(2,0),c(2,1),c(2,2)), # backwards "L"
list(c(0,0),c(0,1),c(0,2),c(0,3)), # |
list(c(0,0),c(0,1),c(1,0),c(1,1))) # small square
```




## Part 1
I'm not really sure how to do this.


```{r}

baseshaft<-matrix(NA,ncol=7,nrow=5*10000000)
dropshapes<-function(shaft,dirlist,counter){
  d<-1
  ### highest point is 0, for the floor
  highestpoint<-0
  cnt<-1
  while(cnt<=counter){
    ### looking for patterns
    #if(cnt%%5==1){cat(cnt,d,highestpoint,"\n")} ### this is to help gather data for part 2
    ### spawn a shape at the right place
    nowshape<-shapes[[((cnt-1)%%5)+1]]
    ###start with three blank rows (or 4 above the highest point)
    ###and two blank columns, so the shapes "0" goes to 3
    nowshape<-lapply(nowshape,function(x){x+c(3,(highestpoint+4))})
    keepgoing<-TRUE
    while(keepgoing==TRUE){
      lr<-dirlist[d]
      if(lr==("<")){
        moveleft<-lapply(nowshape,function(x){x-c(1,0)})
        ## check through maybe.  if any are 0 then false.
        ## if any are in spaces that are already occupied then false
        ## if all are true, then shape<-maybe
        chk<-sapply(moveleft,function(x){
          if(x[1]==0){FALSE
          }else if(!is.na(shaft[x[2],x[1]])){FALSE}else{TRUE}})
        if(all(chk)){nowshape<-moveleft}
      }else{
        moveright<-lapply(nowshape,function(x){x+c(1,0)})
        ## check through maybe.  if any are 8 then false.
        ## if any are in spaces that are already occupied then false
        ## if all are true, then nowshape<-maybe
        chk<-sapply(moveright,function(x){
          if(x[1]==8){FALSE
          }else if(!is.na(shaft[x[2],x[1]])){FALSE}else{TRUE}})
        if(all(chk)){nowshape<-moveright}
      }### try to drop the shape
      allclear<-sapply(nowshape,function(x){
        ## if on the floor then FALSE
        if(x[2]==1){FALSE
        }else{
          ### if underneath is space, then TRUE, otherwise FALSE
          x<-x-c(0,1)
          if(is.na(shaft[x[2],x[1]])){TRUE}else{FALSE}}})
      if(all(allclear)){
        ### if everything below is clear, drop the shape down one
        ### get the next direction
        ### keep going
        nowshape<-lapply(nowshape,function(x){x-c(0,1)})
        d<-d+1
        if(d>length(dirlist)){d<-1}
      }else{
        ### add to the bottom of the shaft
        for(p in 1:length(nowshape)){
          shaft[nowshape[[p]][2],nowshape[[p]][1]]<-cnt}
        ### find the highest point
        nhigh<-max(sapply(nowshape,function(x)x[2]))
        highestpoint<-max(highestpoint,nhigh)
        d<-d+1
        if(d>length(dirlist)){d<-1}
        keepgoing<-FALSE}}
      cnt<-cnt+1}
    list(highestpoint,shaft)}
```


```{r}
part1<-dropshapes(baseshaft,jets,2022)[[1]]
part1
```

##Part 2

As, obviously, there's no way to drop a trillion of these, there has to be a pattern of some sort.

For my real data - in excel, I can figure out that every 1700 stones the highest point grows 2600.
(this was through looking at the highest point for the beginning of each cycle and seeing that it repeated every every 340 cycles (or 1700 stones))

The cycle for the sample data is different, the highest point increases 53 every 35 stones

```{r}
## real data is commented out
##1000000000000%/%1700
##1000000000000%%1700

1000000000000%/%35
1000000000000%%35
```


```{r}
#So, if I find out where it is at 1900 (it takes a while to balance out, so start from 1900 rather than 200), then 
##take that, add 588235293*2660, it should be right.
##basecycle<-dropshapes(baseshaft,jets,1900)[[1]]
##part2<-basecycle+(588235293*2660)
##part2

basecycle<-dropshapes(baseshaft,jets,50)[[1]]
part2<-basecycle+(28571428570*53)
part2
```
