---
title: "Day 24 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
conflicts_prefer(dplyr::filter)
library(stringr)
library(tidyverse)
library(ggraph)
library(readr)
library(collections)
options(scipen = 999)
```


## Part 1

```{r}
input<-read_lines("Day24Sample2.txt")
blizzardmap<-matrix(ncol=nchar(input[1]),nrow=length(input))
for(i in 1:length(input)){
  for(j in 1:nchar(input[1])){
    blizzardmap[i,j]<-str_sub(input[i],j,j)}}
```

Run this as a search.

Basically, it will pull something off the queue, check up, down, left, right, and stay in place to see if there's a going to be a blizzard next step. If not put that move in the queue (with time  & priority based on manhattan distance & time)

How to know if there's a blizzard? Multiple dictionaries - because the blizzards are cyclical, they can be pre-mapped.

```{r}
mazerunner<-function(sp=c(2,1),ep,st=0,bm=blizzardmap){
  ### start by setting up the maps
  hm<-dict()  #horizontal dictionary for < & > - repeats every bmwidth  
  vm<-dict()  #vertical dictionary for v & ^ - repeats every bmheight
  bmwidth<-ncol(bm)-2 #ignore the edges
  bmheight<-nrow(bm)-2   #ignore top bottom edges
  for(x in 1:ncol(bm)){
    for(y in 1:nrow(bm)){
      s<-bm[y,x]
      if(s %in% c("#","v","^","<",">")){
        switch(s,
               #  #s are stable, appear in every lookup
               "#"={for(v in 0:(bmheight-1)){
                 vm$set(toString(c(v,x,y)),"#")} 
                 for(h in 0:(bmwidth-1)){
                   hm$set(toString(c(h,x,y)),"#")}},
                 # v's go into the vm dict - 
                 "v"={for(v in 0:(bmheight-1)){
                   flipy<-if(y+v>=nrow(bm)){y+v-bmheight}else{y+v}
                   vm$set(toString(c(v,x,flipy)),"v")}},
                 "^"={for(v in 0:(bmheight-1)){
                   flipy<-if(y-v<=1){y-v+bmheight}else{y-v}
                   vm$set(toString(c(v,x,flipy)),"^")}},
                 ">"={for(h in 0:(bmwidth-1)){
                   flipx<-if(x+h>=ncol(bm)){x+h-bmwidth}else{x+h}
                   hm$set(toString(c(h,flipx,y)),">")}},
                 "<"={for(h in 0:(bmwidth-1)){
                   flipx<-if(x-h<=1){x-h+bmwidth}else{x-h}
                   hm$set(toString(c(h,flipx,y)),"<")}},
                 cat("something went wrong\n"))}}}
  #### set up the queue & start
  pq<-priority_queue()
  pq$push(c(st,sp),priority=sum(abs(ep-sp)))
  beenthere<-dict()
  counter<-1
  maxdist<-0
  while(pq$size()>0){
    working<-pq$pop()
    if(beenthere$has(toString(working))){next}
    beenthere$set(toString(working),working)
    if(all(working[2:3]==ep)){
      return(working)
      break}else{
        possibilities<-
          as.data.frame(rbind(c(1,0,0)+working,
                              c(1,1,0)+working,
                              c(1,0,1)+working,
                              c(1,-1,0)+working,
                              c(1,0,-1)+working))
        colnames(possibilities)<-c("t","x","y")
        possibilities<-possibilities %>%rowwise%>%
          filter(x>=1,y>=1,x<=ncol(bm),y<=nrow(bm)) %>%
          filter(!vm$has(toString(c(t%%bmheight,x,y))))%>%
          filter(!hm$has(toString(c(t%%bmwidth,x,y))))%>%
          filter(!beenthere$has(toString(c(t,x,y))))%>%
          mutate(pri=t+2*sum(abs(ep-c(x,y))))
        if(nrow(possibilities>0)){
          for(i in 1:nrow(possibilities)){
            p<-possibilities[i,]
            pq$push(c(p$t,p$x,p$y),priority=-p$pri)}}}
        counter<-counter+1}}
```


```{r}
part1<-mazerunner(c(2,1),c(7,6),0,blizzardmap)
part1
part1<-part1[1]
part1
```




## Part 2 - 
Because "wait" is an option, the fastest to the end can be the start of the fastest back (if one of the later times is a better time to leave, the "wait" option will find it.)

```{r}
backtostart<-mazerunner(c(7,6),c(2,1),part1,blizzardmap)
backtostart<-backtostart[1]
part2<-mazerunner(c(2,1),c(7,6),backtostart,blizzardmap)
part2<-part2[1]
backtostart;part2
```

