---
title: "Day 15 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(igraph)
library(ggplot2)
library(rgl)
library(ggraph)
library(readr)
options(scipen = 999)
```


## Part 1
First put everything into a data frame
```{r}
input<-read_lines("Day15Sample.txt")

sensors<-as.data.frame(matrix(ncol=4,nrow=0))
colnames(sensors)<-c("sx","sy","bx","by")
for(i in 1:length(input)){
  ln<-unlist(str_split(input[i],"[=,:]"))
  ln<-as.numeric(ln[c(2,4,6,8)])
  sensors<-rbind(sensors,list(sx=ln[1],sy=ln[2],bx=ln[3],by=ln[4]))}

```
then find the manhattan distances for each  sensor/beacon pair
```{r}
sensors <- sensors %>% rowwise %>% mutate(md=(abs(bx-sx)+abs(by-sy)))
```

for each sensor, find if its no beacon range even meets the target.
If it does, then, find its "no beacon" range

```{r}
target<-10
possibles<-c()

### find possible places for no beacons
for(i in 1:nrow(sensors)){
  ### how much distance is left after getting to that line
  rmn<-sensors$md[i]-abs(sensors$sy[i]-target)
  if(rmn>=0){
    possibles<-union(possibles,(sensors$sx[i]-rmn):(sensors$sx[i]+rmn))}}
### find beacons on the target line
beaconsattarget<-sensors$bx[which(sensors$by==target)]

### pull those from the possibles, because, they're obviously not possible
possibles<-possibles[-c(which(possibles %in% beaconsattarget))]
part1<-length(possibles)
part1
```
## Part 2
Brute force would be an option, but it would take weeks & I have things to do.

I stole this answer - I was getting nowhere slowly.

The beacon lies at md+1 for some sensors.
First find the md+1 line for all the sensors
Then, find the intersections of all of those lines (at most 4s^2). Then filter those points to the range (0<=x,y<=4000000) the remaining points can be pretty quickly checked.

find the Cs for (Ax+BY=C)
```{r}
sensors<-sensors %>% rowwise %>%
  mutate(aplusb1=sx+(sy+md+1),aplusb2=sx+(sy-md-1),aminusb1=sx-(sy+md+1),aminusb2=sx-(sy-md-1))

aplusbs<-unique(c(sensors$aplusb1,sensors$aplusb2))
aminusbs<-unique(c(sensors$aminusb1,sensors$aminusb2))
```
find solutions:
```{r}
beaconpoints<-as.data.frame(matrix(ncol=2,nrow=0))
for(i in 1:length(aplusbs)){
  for(j in 1:length(aminusbs)){
    possbeacon<-solve(rbind(c(1,1),c(1,-1)),c(aplusbs[i],aminusbs[j]))
    beaconpoints<-rbind(beaconpoints,possbeacon)}}
colnames(beaconpoints)<-c("x","y")
beaconpoints$x<-as.numeric(beaconpoints$x)
beaconpoints$y<-as.numeric(beaconpoints$y)
beaconpoints<-beaconpoints %>% ungroup %>% filter(x>=0,y>=0,x<=20,y<=20)
```
For each point, check to see if it is further away from the sensor than md.  If so, that's the distress beacon.

```{r}
part2<-list()
for(i in 1:nrow(beaconpoints)){
  checkpoint<-beaconpoints[i,]
  checkdistances<-sapply(1:nrow(sensors),function(x){
    if((abs(sensors$sx[x]-checkpoint[1])+abs(sensors$sy[x]-checkpoint[2]))>sensors$md[x]){TRUE}else{FALSE}})
  if(all(checkdistances)){part2<-append(part2,checkpoint)}}

part2
```
and to find the answer:
```{r}
part2<-part2$x*4000000+part2$y
part2
```


