---
title: "Day 14 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(igraph)
library(ggplot2)
library(rgl)
library(ggraph)
library(readr)
library(ds4psy)
library(numbers)
library(reshape2)
options(scipen = 999)
```


## Part 1
Start by building the walls

```{r}
input<-read_lines("Day14Sample.txt")

rocks<-as.data.frame(matrix(ncol=4,nrow=0))
colnames(rocks)<-c("x1","y1","x2","y2")
for(i in 1:length(input)){
  x<-as.numeric(unlist(str_split(str_replace(input[i]," ",""),"[->,]")))
  for(j in seq(from=1,to=(length(x)-2),by=3)){
    rocks<-rbind(rocks,list(x1=x[j],y1=x[j+1],x2=x[j+3],y2=x[j+4]))}}
```

So there's less math, (hah), first, drop y by 1 because the matrices start at 1 rather than 0.
And then because I know part 2, the matrix will be two more than the highest (new) y.
And then, make it a little bit wider than it is tall on the right side.
```{r}
rocks$y1<-rocks$y1+1
rocks$y2<-rocks$y2+1

mxy<-max(max(rocks$y1),max(rocks$y2))
sandpit<-matrix(NA,ncol=500+mxy+4,nrow=mxy+2)
```

Then the rocks go into the pit

```{r}

for(i in 1:nrow(rocks)){
  rockwall<-rocks[i,]
  sandpit[rockwall$y1:rockwall$y2,rockwall$x1:rockwall$x2]<-"R"}
### create a backup for part 2
spbk<-sandpit

```

finally - drop some sand:
```{r}


sandcount<-0
newsand<-c(500,1)

while(newsand[2]<nrow(sandpit)){
  ### if it hits the bottom of the matrix, go to the next sand
  while(newsand[2]<nrow(sandpit)){
    ### if there's nothing below, drop one, keep going
    if(is.na(sandpit[newsand[2]+1,newsand[1]])){
      newsand[2]<-newsand[2]+1
      next
      ### if there's something below, but nothing to the left, 
    }else if(is.na(sandpit[newsand[2]+1,newsand[1]-1])){
      newsand[2]<-newsand[2]+1
      newsand[1]<-newsand[1]-1
      next
      ### if there's something below & to the left, check the right
    }else if(is.na(sandpit[newsand[2]+1,newsand[1]+1])){
      newsand[2]<-newsand[2]+1
      newsand[1]<-newsand[1]+1
      next
      ### else, the sand rests and start with the next grain
    }else{sandpit[newsand[2],newsand[1]]<-"S"
    sandcount<-sandcount+1
    newsand<-c(500,1)
    break}}}

part1<-sandcount
part1
```
## Part 2

For part 2, put a floor down.  This time, stop when there's a grain of sand at 500,1
```{r}

### restore from the backup
sandpit<-spbk
### add the floor
sandpit[mxy+2,]<-"R"
sandcount<-0
newsand<-c(500,1)
while(is.na(sandpit[1,500])){
  ### if it hits the bottom of the matrix, go to the next sand
  while(is.na(sandpit[1,500])){
    ### if there's nothing below, drop one, keep going
    if(is.na(sandpit[newsand[2]+1,newsand[1]])){
      newsand[2]<-newsand[2]+1
      next
      ### if there's something below, but nothing to the left, 
    }else if(is.na(sandpit[newsand[2]+1,newsand[1]-1])){
      newsand[2]<-newsand[2]+1
      newsand[1]<-newsand[1]-1
      next
      ### if there's something below & to the left, check the right
    }else if(is.na(sandpit[newsand[2]+1,newsand[1]+1])){
      newsand[2]<-newsand[2]+1
      newsand[1]<-newsand[1]+1
      next
      ### otherwise, add a grain of sand, and reset for the next
    }else{sandpit[newsand[2],newsand[1]]<-"S"
    sandcount<-sandcount+1
    newsand<-c(500,1)
    break}}}

part2<-sandcount
part2
```


```{r,echo=FALSE,warning=FALSE,message=FALSE}
### for the graph
input<-read_lines("../../AoCData/AOC2022/Day14.txt")

### redo the input & rocks
rocks<-as.data.frame(matrix(ncol=4,nrow=0))
colnames(rocks)<-c("x1","y1","x2","y2")
for(i in 1:length(input)){
  x<-as.numeric(unlist(str_split(str_replace(input[i]," ",""),"[->,]")))
  for(j in seq(from=1,to=(length(x)-2),by=3)){
    rocks<-rbind(rocks,list(x1=x[j],y1=x[j+1],x2=x[j+3],y2=x[j+4]))}}
rocks$y1<-rocks$y1+1
rocks$y2<-rocks$y2+1
### recreate the sandpit
mxy<-max(max(rocks$y1),max(rocks$y2))
sandpit<-matrix(NA,ncol=500+mxy+4,nrow=mxy+2)
for(i in 1:nrow(rocks)){
  rockwall<-rocks[i,]
  sandpit[rockwall$y1:rockwall$y2,rockwall$x1:rockwall$x2]<-"R"}
### create a backup for part 2
spbk<-sandpit

#### rerun the sand
sandpit[mxy+2,]<-"R"
sandcount<-0
newsand<-c(500,1)
while(is.na(sandpit[1,500])){
  ### if it hits the bottom of the matrix, go to the next sand
  while(is.na(sandpit[1,500])){
    ### if there's nothing below, drop one, keep going
    if(is.na(sandpit[newsand[2]+1,newsand[1]])){
      newsand[2]<-newsand[2]+1
      next
      ### if there's something below, but nothing to the left, 
    }else if(is.na(sandpit[newsand[2]+1,newsand[1]-1])){
      newsand[2]<-newsand[2]+1
      newsand[1]<-newsand[1]-1
      next
      ### if there's something below & to the left, check the right
    }else if(is.na(sandpit[newsand[2]+1,newsand[1]+1])){
      newsand[2]<-newsand[2]+1
      newsand[1]<-newsand[1]+1
      next
    }else{sandpit[newsand[2],newsand[1]]<-"S"
    sandcount<-sandcount+1
    newsand<-c(500,1)
    break}}}

bkpit<-sandpit

### find the min & max of the sandpit
lowx<-min(which(sandpit[(nrow(sandpit)-1),]=="S"))
hix<-max(which(sandpit[(nrow(sandpit)-1),]=="S"))
sandpit<-sandpit[,(lowx-2):(hix+2)]

meltsand<-melt(sandpit)



ggplot(meltsand,aes(x=Var2,y=Var1, colour=value))+geom_point()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position="none")+
  scale_y_reverse()+
  coord_fixed()
```
