---
title: "Day 22 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(igraph)
library(ggplot2)
library(rgl)
library(ggraph)
library(readr)
library(ds4psy)
library(numbers)
options(scipen = 999)
```


## Part 1
Read in data - break in two: the map & the directions
```{r}
input<-read_lines("Day22Sample.txt")

blnk<-which(input=="")
route<-input[(blnk+1)]
boardmap<-matrix(NA,ncol=max(nchar(input)),nrow=(length(input)-2))
for(i in 1:(length(input)-2)){
  for(j in 1:nchar(input[i])){
    if(str_sub(input[i],j,j) !=" "){boardmap[i,j]<-str_sub(input[i],j,j)}}}

```

Find the path

Process:
1. Start with a position and direction:
2. Find how many steps you're going to take & where you turn at the end of your walk
3. Figure out where your next step *should* be - 
3.a either it's right in front of you, or
3.b it's around the edge
4. Find out if there's a rock there:
  4.a if there's a rock, stop moving
  4.b if there's no rock, step there and go back to step 3 if you have more steps
5. Make this your new starting point
6. Find your new direction based on your turn
  6.a if you have a new direction, go back to step 2.
  6.b if you are at the end of the walk, stop - report where you are and where you're facing.


```{r}
quickpath<-function(rte=route,bm=boardmap){
  counter<-1
  
  ### find the starting point
  currp<-c((min(which(!(is.na(bm[1,]))))),1)
  currd<-"E"
  while(!is.na(rte)){
    ### pull the first place from the routelist
    cutoff<-str_locate(rte,"[RL]")[1]
    if(is.na(cutoff)){currrte<-rte}else{currrte<-str_sub(rte,1,cutoff)}
    rte<-str_sub(rte,cutoff+1)
    if(str_sub(currrte,-1)%in%(c("R","L"))){
      stps<-as.numeric(str_sub(currrte,end=-2))
      nxtd<-str_sub(currrte,-1)}else{stps<-as.numeric(currrte)
      nxtd<-"stop"}
    ### take steps
    switch(currd,
           "N"={s<-1
           while(s<=stps){
             #find next step (wrap around if needed)
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             #find next step (wrap around if needed)
             nxtx<-currp[1]
             nxty<-currp[2]-1
             if(nxty<1||is.na(bm[nxty,nxtx])){nxty<-max(which(!is.na(bm[,nxtx])))}
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             if(bm[nxty,nxtx]=="#"){break}else{currp<-c(nxtx,nxty)}
             s<-s+1}
           if(nxtd=="L"){currd<-"W"}else if(nxtd=="R"){currd<-"E"}},
           "E"={s<-1
           while(s<=stps){
             #find next step (wrap around if needed)
             nxtx<-currp[1]+1
             nxty<-currp[2]
             if(nxtx>ncol(bm)||is.na(bm[nxty,nxtx])){nxtx<-min(which(!is.na(bm[nxty,])))}
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             if(bm[nxty,nxtx]=="#"){break}else{currp<-c(nxtx,nxty)}
             s<-s+1}
           if(nxtd=="L"){currd<-"N"}else if(nxtd=="R"){currd<-"S"}},
           "S"={s<-1
           while(s<=stps){
             #find next step (wrap around if needed)
             nxtx<-currp[1]
             nxty<-currp[2]+1
             if(nxty>nrow(bm)||is.na(bm[nxty,nxtx])){nxty<-min(which(!is.na(bm[,nxtx])))}
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             if(bm[nxty,nxtx]=="#"){break}else{currp<-c(nxtx,nxty)}
             s<-s+1}
           if(nxtd=="L"){currd<-"E"}else if(nxtd=="R") {currd<-"W"}},
           "W"={s<-1
           while(s<=stps){
             #find next step (wrap around if needed)
             nxtx<-currp[1]-1
             nxty<-currp[2]
             if(nxtx<1||is.na(bm[nxty,nxtx])){nxtx<-max(which(!is.na(bm[nxty,])))}
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             if(bm[nxty,nxtx]=="#"){break}else{currp<-c(nxtx,nxty)}
             s<-s+1}
           if(nxtd=="L"){currd<-"S"}else if(nxtd=="R"){currd<-"N"}},
           cat("no direction\n"))
    ### end of while
    counter<-counter+1
  }### outside of the while
  cat(counter,currp,currd,"\n")
  psswd<-
    switch(currd,
           "N"=3,
           "W"=2,
           "E"=0,
           "S"=1,
           1000000000)+1000*currp[2]+4*currp[1]
  
  psswd}
```

```{r}
### quick stuff


```


```{r}
part1<-quickpath()
part1

# 46 162 W 
# [1] 162186

```
### Part 2

First, doing this with just the flat map to make sure the idea works - but there's a slight change in process:

Process:
1. Start with a position and direction:
2. Find how many steps you're going to take & where you turn at the end of your walk (*tell the world you are not falling off of a cliff*)
3. Figure out where your next step *should* be - 
  3.a either it's right in front of you, or
  3.b *you are about to fall off a cliff* - let everyone know about the cliff & wraparound to find the next step.
4. Find out if there's a rock there:
  4.a if there's a rock, stop moving - *and let everyone know you are not falling off a cliff* whether or not that was a possibility.
  4.b if there's no rock, step there
    4.b.i if you did not fall off a cliff, go back to step 3 if you have more steps
    4.b.ii *reattach the remainder of your steps & the turn to the route list* you are not taking any more steps until the whole cliff thing is worked out.
5. Make this your new starting point (either way)
6. 
  6.a *If you did not fall off of a cliff* Find your new direction based on your turn - go back to step 2.
  6.b *if you did fall off a cliff* keep your current direction - go to step 2 for the rest of the steps before the cliff fall
  6.c if you are at the end of the walk, stop - report where you are and where you're facing.

```{r}
flatpath<-function(rte=route,bm=boardmap){
  counter<-1
  ### find the starting point
  currp<-c((min(which(!(is.na(bm[1,]))))),1)
  currd<-"E"
  while(!is.na(rte)){
    ### pull the first place from the routelist
    cutoff<-str_locate(rte,"[RL]")[1]
    if(is.na(cutoff)){currrte<-rte}else{currrte<-str_sub(rte,1,cutoff)}
    rte<-str_sub(rte,cutoff+1)
    regroup<-FALSE
    if(str_sub(currrte,-1)%in%(c("R","L"))){
      stps<-as.numeric(str_sub(currrte,end=-2))
      nxtd<-str_sub(currrte,-1)}else{stps<-as.numeric(currrte)
      nxtd<-"stop"}
    ### take steps
    switch(currd,
           "N"={s<-1
           while(s<=stps){
             #find next step (wrap around if needed)
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             #find next step (wrap around if needed)
             nxtx<-currp[1]
             nxty<-currp[2]-1
             if(nxty<1||is.na(bm[nxty,nxtx])){nxty<-max(which(!is.na(bm[,nxtx])))
             ### add fla for regroup
             regroup<-TRUE}
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             if(bm[nxty,nxtx]=="#"){
               regroup<-FALSE
               break}else{currp<-c(nxtx,nxty)}
             if(regroup){rte<-str_c(stps-s,nxtd,rte)
             break}
             s<-s+1}
           if(!regroup){if(nxtd=="L"){currd<-"W"}else if(nxtd=="R"){currd<-"E"}}},
           "E"={s<-1
           while(s<=stps){
             #find next step (wrap around if needed)
             nxtx<-currp[1]+1
             nxty<-currp[2]
             ### wrap around position
             if(nxtx>ncol(bm)||is.na(bm[nxty,nxtx])){nxtx<-min(which(!is.na(bm[nxty,])))
             ####Add a flag to say that we should stop & start if we have to drive off the cliff. 
             regroup<-TRUE}
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             if(bm[nxty,nxtx]=="#"){
               regroup<-FALSE
               break}else{
                 currp<-c(nxtx,nxty)}
             ### add the lines to 
             if(regroup){rte<-str_c(stps-s,nxtd,rte)
             break}
             s<-s+1}
           #### make sure that we don't change things if there's a regroup
           if(!regroup){if(nxtd=="L"){currd<-"N"}else if(nxtd=="R"){currd<-"S"}}},
           "S"={s<-1
           while(s<=stps){
             #find next step (wrap around if needed)
             nxtx<-currp[1]
             nxty<-currp[2]+1
             if(nxty>nrow(bm)||is.na(bm[nxty,nxtx])){nxty<-min(which(!is.na(bm[,nxtx])))
             ###regroup
             regroup<-TRUE}
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             if(bm[nxty,nxtx]=="#"){
               regroup<-FALSE
               break}else{currp<-c(nxtx,nxty)}
             if(regroup){rte<-str_c(stps-s,nxtd,rte)
             break}
             s<-s+1}
           if(!regroup){if(nxtd=="L"){currd<-"E"}else if(nxtd=="R") {currd<-"W"}}},
           "W"={s<-1
           while(s<=stps){
             #find next step (wrap around if needed)
             nxtx<-currp[1]-1
             nxty<-currp[2]
             if(nxtx<1||is.na(bm[nxty,nxtx])){nxtx<-max(which(!is.na(bm[nxty,])))
             regroup<-TRUE}
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             if(bm[nxty,nxtx]=="#"){
               regroup<-FALSE
               break}else{currp<-c(nxtx,nxty)}
             if(regroup){rte<-str_c(stps-s,nxtd,rte)
             break}
             s<-s+1}
           if(!regroup){if(nxtd=="L"){currd<-"S"}else if(nxtd=="R"){currd<-"N"}}},
           cat("no direction\n"))
    ### end of while
    counter<-counter+1
  }### outside of the while
  cat(counter,currp,currd,"\n")
  psswd<-
    switch(currd,
           "N"=3,
           "W"=2,
           "E"=0,
           "S"=1,
           1000000000)+1000*currp[2]+4*currp[1]
  
  psswd}
```



```{r}
part1<-flatpath()
part1

```
Now do the same thing with the cube - (this is not a general solution, it is specific for the shape of the input.)


Process:
1. Start with a position and direction:
2. Find how many steps you're going to take & where you turn at the end of your walk (*tell the world you are not falling off of a cliff*)
3. Figure out where your next step *should* be - 
  3.a either it's right in front of you, or
  3.b *you are about to fall off a cliff* - let everyone know about the cliff & **follow some extremely complicated instructions to find your next step based on your current & next position**
4. Find out if there's a rock there:
  4.a if there's a rock, stop moving - *and let everyone know you are not falling off a cliff* whether or not that was a possibility.
  4.b if there's no rock, step there
    4.b.i if you did not fall off a cliff, go back to step 3 if you have more steps
    4.b.ii *reattach the remainder of your steps & the turn to the route list* you are not taking any more steps until the whole cliff thing is worked out.
5. Make this your new starting point (either way)
6. 
  6.a *If you did not fall off of a cliff* Find your new direction based on your turn - go back to step 2.
  6.b *if you did fall off a cliff* **again, follow some complicated instructions to find your next direction** - go to step 2 for the rest of the steps before the cliff fall
  6.c if you are at the end of the walk, stop - report where you are and where you're facing.


direction is based on the flat map.

And what are the complicated instructions?
If you were heading North:
a. for 1<=x<=50:    head E - x=51, y=old x+50 (opposite m)
b. for 51=x<=100:   head E - x=1, y=old x+100 (opposite o)
c. for 101<=x<=150: head N - x=old x - 100, y=200   (opposite h)

If you were heading East:
d. for 1<=y<=50:    head W - x=1, y=151-old y (opposite f)
e. for 51=y<=100:   head N - x=oldy+50,   y=50 (opposite j)
f. for 101<=y<=150: head W - x=150 y=  y=151-old y  (opposite d) 
g. for 151<=y<=200: head N - y=150, x=oldy-100  (opposite i)

If you were heading South:
h. for 1<=x<=50:     head S - y=1, x=oldx+100 (opposite c)
i. for 51=x<=100:    head W, x=150, y=oldx+100 (opposite g)
j. for 101<=x<=150:  head W, x=100, y=oldx-50 (opposite e) 

If you were heading West:
l. for 1<=y<=50:   head E - x=1, y=151-oldy (opposite n)
m. for 51=y<=100:  head S - x=oldy-50, y=101 (opposite a) 
n. for 101<=y<=150:head E - x=51, y=151-oldy (opposite l)
o. for 151<=y<=200:head S - x=oldy-100, y=1   (opposite b) 


```{r}
cubepath<-function(rte=route,bm=boardmap){
  counter<-1
  
  ### find the starting point
  currp<-c((min(which(!(is.na(bm[1,]))))),1)
  currd<-"E"
  while(!is.na(rte)){
    ### pull the first place from the routelist
    cutoff<-str_locate(rte,"[RL]")[1]
    if(is.na(cutoff)){currrte<-rte}else{currrte<-str_sub(rte,1,cutoff)}
    rte<-str_sub(rte,cutoff+1)
    regroup<-FALSE
    if(str_sub(currrte,-1)%in%(c("R","L"))){
      stps<-as.numeric(str_sub(currrte,end=-2))
      nxtd<-str_sub(currrte,-1)}else{stps<-as.numeric(currrte)
      nxtd<-"stop"}
    ### take steps
    switch(currd,
           "N"={s<-1
           while(s<=stps){
             #find next step (wrap around if needed)
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             nxtx<-currp[1]
             nxty<-currp[2]-1
             if(nxty<1||is.na(bm[nxty,nxtx])){
               if(currp[1]<=50){
                 nxtx<-51
                 nxty<-currp[1]+50
                 cliffd<-"E"
               }else if(currp[1]>=51&&currp[1]<=100){
                 nxtx<-1
                 nxty<-currp[1]+100
                 cliffd<-"E"
               } else{
                 nxtx<-currp[1]-100
                 nxty<-200
                 cliffd<-"N"}
               ### add flag for regroup
               regroup<-TRUE}
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             if(bm[nxty,nxtx]=="#"){
               regroup<-FALSE
               break}else{currp<-c(nxtx,nxty)}
             if(regroup){rte<-str_c(stps-s,nxtd,rte)
             break}
             s<-s+1}
           if(regroup){currd<-cliffd
           } else {if(nxtd=="L"){currd<-"W"}else if(nxtd=="R"){currd<-"E"}}},
           "E"={s<-1
           while(s<=stps){
             #find next step (wrap around if needed)
             nxtx<-currp[1]+1
             nxty<-currp[2]
             ### cliff position & direction
             if(nxtx>ncol(bm)||is.na(bm[nxty,nxtx])){
               if(currp[2]<=50){
                 nxtx<-100
                 nxty<-151-currp[2]
                 cliffd<-"W"
               }else if(currp[2]>=51&&currp[2]<=100){
                 nxtx<-currp[2]+50
                 nxty<-50
                 cliffd<-"N"
               } else if(currp[2]>=101&&currp[2]<=150){
                 nxtx<-150
                 nxty<-151-currp[2]
                 cliffd<-"W"
               }else{
                 nxtx<-currp[2]-100
                 nxty<-150
                 cliffd<-"N"} 
               ####Add a flag to say that we should stop & start if we have to drive off the cliff. 
               regroup<-TRUE}
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             if(bm[nxty,nxtx]=="#"){
               regroup<-FALSE
               break}else{
                 currp<-c(nxtx,nxty)}
             ### add the lines to 
             if(regroup){rte<-str_c(stps-s,nxtd,rte)
             break}
             s<-s+1}
           #### if not at stop, change directions.
           if(regroup){currd<-cliffd
           } else {if(nxtd=="L"){currd<-"N"}else if(nxtd=="R"){currd<-"S"}}},
           "S"={s<-1
           while(s<=stps){
             #find next step (wrap around if needed)
             nxtx<-currp[1]
             nxty<-currp[2]+1
             if(nxty>nrow(bm)||is.na(bm[nxty,nxtx])){
               if(currp[1]<=50){
                 nxtx<-currp[1]+100
                 nxty<-1
                 cliffd<-"S"
               }else if(currp[1]>=51&&currp[1]<=100){
                 nxtx<-50
                 nxty<-currp[1]+100
                 cliffd<-"W"
               } else{
                 nxtx<-100
                 nxty<-currp[1]-50
                 cliffd<-"W"}           
               ###regroup
               regroup<-TRUE}
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             if(bm[nxty,nxtx]=="#"){
               regroup<-FALSE
               break}else{currp<-c(nxtx,nxty)}
             if(regroup){rte<-str_c(stps-s,nxtd,rte)
             break}
             s<-s+1}
           if(regroup){currd<-cliffd
           } else {if(nxtd=="L"){currd<-"E"}else if(nxtd=="R"){currd<-"W"}}},
           "W"={s<-1
           while(s<=stps){
             #find next step (wrap around if needed)
             nxtx<-currp[1]-1
             nxty<-currp[2]
             ### cliff position & direction
             if(nxtx<1||is.na(bm[nxty,nxtx])){
               if(currp[2]<=50){
                 nxtx<-1
                 nxty<-151-currp[2]
                 cliffd<-"E"
               }else if(currp[2]>=51&&currp[2]<=100){
                 nxtx<-currp[2]-50
                 nxty<-101
                 cliffd<-"S"
               } else if(currp[2]>=101&&currp[2]<=150){
                 nxtx<-51
                 nxty<-151-currp[2]
                 cliffd<-"E"
               }else{
                 nxtx<-currp[2]-100
                 nxty<-1
                 cliffd<-"S"}  
               regroup<-TRUE}
             #find if it's a #
             #if it is clear, move there.
             #if it is blocked stay where I am & break
             if(bm[nxty,nxtx]=="#"){
               regroup<-FALSE
               break}else{currp<-c(nxtx,nxty)}
             if(regroup){rte<-str_c(stps-s,nxtd,rte)
             break}
             s<-s+1}
           if(regroup){currd<-cliffd
           } else {if(nxtd=="L"){currd<-"S"}else if(nxtd=="R"){currd<-"N"}}},
           cat("no direction\n"))
    ### end of while
    counter<-counter+1
  }### outside of the while
  cat(counter, currp,currd,"\n")
  psswd<-
    switch(currd,
           "N"=3,
           "W"=2,
           "E"=0,
           "S"=1,
           1000000000)+1000*currp[2]+4*currp[1]
  
  psswd}
```

Because the sample is a different shape an dimensions to the solution, it won't work.

```{r,eval=FALSE,message=FALSE}
part2<-cubepath()
part2
```
