---
title: "Day 16 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(collections)
library(reshape2)
library(stringr)
library(tidyverse)
library(igraph)
library(ggplot2)
library(rgl)
library(ggraph)
library(readr)
library(ds4psy)
library(numbers)
options(scipen = 999)
```


```{r}
input<-read_lines("Day16Sample.txt")
```

## Part 1
Start by reformating the input
```{r}
valvelist<-lapply(input,function(x){
  x<-str_replace(x,"Valve ","")
  x<-str_replace(x," has flow rate="," ")
  x<-str_replace(x,"; tunnels lead to valves "," ")
  x<-str_replace(x,"; tunnel leads to valve "," ")
  x<-str_replace_all(x,", ",",")
  x<-unlist(str_split(x," "))
  x})
```

In the interest of making this a little less onerous, make this graph a little smaller by removing the broken valves (no need to stop there), except for the starting point (AA)

```{r}
### create the dictionary for the valves 
flowdict<-dict()
for(i in 1:length(valvelist)){
  flowdict$set(valvelist[[i]][1],as.numeric(valvelist[[i]][2]))}

### create the large map
### edgelist
el<-as.data.frame(matrix(ncol=3,nrow=0))
### list of valves for later - need AA & the working valves
workingvalves<-c("AA")
for(i in 1:length(valvelist)){
  f<-valvelist[[i]][1]
  t<-unlist(str_split(valvelist[[i]][3],","))
  n<-as.numeric(valvelist[[i]][2])
  for(j in 1:length(t)){
    el<-rbind(el,list(f=f,t=t[j],d=n))}
  if(n>0){workingvalves<-c(workingvalves,f)}}

valvemap<-simplify(graph_from_edgelist(as.matrix(el[,1:2]),directed=FALSE))

### a quick graph to see what this looks like
ggraph(valvemap,layout="stress")+
  geom_node_text(aes(label=name))+
  geom_edge_fan(start_cap = circle(3, 'mm'),
                end_cap = circle(3, 'mm'))

```
find the abbreviated graph & distances between all of the nodes with values & AA
```{r}
###get the meaningful distances
ds<-distances(valvemap,v=workingvalves,to=workingvalves,mode=all)
ds<-melt(ds)
colnames(ds)<-c("f","t","d")
ds$f<-as.character(ds$f)
ds$t<-as.character(ds$t)
### never going back to AA, so remove that.  And add 1 account for the time to turn on the valve
ds<-ds %>% ungroup %>% filter(t!="AA") %>% rowwise %>%
  mutate(flow=flowdict$get(t),d=d+1)

```

From there - search the way through:

```{r}
volcanovalves<-function(maxtime,valveframe){
  valvestack<-stack()
  ## keep track of the current position, visited positions, time left, and flow
  strt<-list("AA",c("AA"),maxtime,0)
  valvestack$push(strt)
  ### will keep a copy of every step along the way - for part 2
  rtn<-list()
  while(valvestack$size()>0){
   curr<-valvestack$pop()
   pos<-curr[[1]]
   pth<-curr[[2]]
   tleft<-as.numeric(curr[[3]])
   fl<-as.numeric(curr[[4]])
   ## if there's no time left, go to the next one

      if(tleft<0){next
   }else{
     ## otherwise, add this to one of the valid outcomes
     rtn<-append(rtn,list(curr))
     ## check to see if there's anywhere else to go, if so, add that to the stack  
     ndf<-valveframe %>% filter(f==pos,!(t%in%pth))
     if(nrow(ndf>0)){
       for(i in 1:nrow(ndf)){
         x<-ndf[i,]
         nxt<-list(x$t,c(pth,x$t),tleft-x$d,fl+(x$flow*(tleft-x$d)))
         valvestack$push(nxt)}}}}
  rtn}
```



```{r}
part1<-volcanovalves(30,ds)
### find the flow from each possibility.  the max is the answer to part 1
part1<-max(sapply(part1,function(x){x[[4]]}))
part1
```
## Part 2

For part 2, first - run with 26 minutes

```{r}
elephant<-volcanovalves(26,ds)
```

then find combinations where two paths do not overlap (except for at "AA") and add their flows.   The max is the answer for part 2.  To get that (relatively) quickly - sort elephant by the flow and start with the combinations.  Don't even bother if the combination will (obviously) be too small
```{r}
elephant<-elephant[order(-sapply(elephant,function(x){x[[4]]}))]
part2<-0
i<-1
while(i<length(elephant)){
  ### if the flow is less than twice the current max flow, no reason to keep going
  if(2*elephant[[i]][[4]]<part2){next
  }else{
    j<-i+1
    while(j<=length(elephant)){
      ## if i+j< current max, no reason to keep going with this i
      if(elephant[[i]][[4]]+elephant[[j]][[4]]<part2){
        i<-i+1
        j<-i+1
        next
      }else{
        ### if there's no overlap between the two groups,
        ### the new part 2 is the combination of their flows.
        if(!(any(elephant[[i]][[2]][-1]%in%elephant[[j]][[2]][-1]))){
          part2<-elephant[[i]][[4]]+elephant[[j]][[4]]}
        j<-j+1}}}
  i<-i+1}

part2
```
