---
title: "Day 9 Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(stringr)
library(tidyverse)
library(readr)
options(scipen = 999)
```

```{r}
input<-read_lines("Day9Sample.txt")
```

## Part 1

First, build out the path for the head of the rope, it will need to be step by step because of the way that the head pulls the tail:

```{r}
headpath<-list(c(0,0))
for(i in 1:length(input)){
  inst<-unlist(str_split(input[i]," "))
  dir<-inst[1];st<-as.numeric(inst[2])
  for(j in 1:st){
    prev<-unlist(tail(headpath,1))
    switch(dir,
           "R"=nxt<-prev+c(1,0),
           "L"=nxt<-prev-c(1,0),
           "U"=nxt<-prev-c(0,1),
           "D"=nxt<-prev+c(0,1),
           cat("something is wrong\n"))
    headpath<-append(headpath,list(c(nxt)))}}
```

Now find the tailpath, for this treat x & y separately.
If the tail is within 1 of the head, don't move.
Otherwise, if the tail is 2 away from the head in the x direction, move in that direction by 1, move y to match the head's y. Do the same for 

```{r}
tailpath<-function(hp){
  tp<-list(c(0,0))
  for(i in 2:length(hp)){
    headpos<-hp[[i]]
    tailpos<-unlist(tail(tp,1))
    d<-abs(headpos-tailpos)
    ### the head is within one of the tail
    if(all(d<=1)){newx<-tailpos[1]
    newy<-tailpos[2]
    ### if the head has moved up to two away in the x direction - but not the y
    }else if(d[1]==2&&d[2]<2){
      newx<-mean(c(headpos[1],tailpos[1]))
      newy<-headpos[2]
    ### if the head has moved up to two away in the y direction, but not the x  
    }else if(d[2]==2&&d[1]<2){
      newy<-mean(c(headpos[2],tailpos[2]))
      newx<-headpos[1]
    ### if the head is two away in both directions  
    }else if(d[1]==2&&d[2]==2){
      newx<-mean(c(headpos[1],tailpos[1]))
      newy<-mean(c(headpos[2],tailpos[2]))
    }else{cat("something went wrong\n")}
    tp<-append(tp,list(c(newx,newy)))}
  tp}
```

```{r}
part1<-tailpath(headpath)
part1<-length(unique(part1))
part1
```
## Part 2
Repeat this with 10 knots (or in other words, 9 times.)  This needs new data

With the original data:
```{r}
lotsknots<-headpath
for(i in 1:9){
  lotsknots<-tailpath(lotsknots)}

part2<-length(unique(lotsknots))
part2  
```
or for a second sample:

```{r}
## longer input
input<-read_lines("Day9Sample2.txt")

## new headpath
p2path<-list(c(0,0))
for(i in 1:length(input)){
  inst<-unlist(str_split(input[i]," "))
  dir<-inst[1];st<-as.numeric(inst[2])
  for(j in 1:st){
    prev<-unlist(tail(p2path,1))
    switch(dir,
           "R"=nxt<-prev+c(1,0),
           "L"=nxt<-prev-c(1,0),
           "U"=nxt<-prev-c(0,1),
           "D"=nxt<-prev+c(0,1),
           cat("something is wrong\n"))
    p2path<-append(p2path,list(c(nxt)))}}

## rerun
for(i in 1:9){
  p2path<-tailpath(p2path)}

part2<-length(unique(p2path))
part2
```

